---
title: "Lab 2 Homework"
author: "Vincent Castillo"
date: "Winter 2020"
output:
  html_document:
    keep_md: TRUE
    theme: darkly
    toc: yes
    toc_float: yes
    code_folding: show 
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run.  

```{r}
library("tidyverse")
```

## Question 1
### From which publication are these data taken from? Don't do an internet search; show the code that you would use to find out in R. 
```{r}
?msleep
```
_V. M. Savage and G. B. West. A quantitative, theoretical framework for understanding mammalian sleep. Proceedings of the National Academy of Sciences, 104 (3):1051-1056, 2007._

## Question 2  
### Put these data into a new object `sleep`. Make sure they are organized as a data frame.  
```{r}
sleep <- data.frame(msleep)
```

## Question 3  
### Show a list of the column names in this data frame.  
```{r}
colnames(sleep)
```

## Question 4  
### Use `glimpse()` to summarize the data in `sleep`.  
```{r}
glimpse(sleep)
```

## Question 5  
### Use `summary()` to summarize the data in `sleep`.  
```{r}
summary(sleep)
```

## Question 6  
### Print out the first column of the data frame and then print out the first row.  
```{r}
sleep[, 1]
```

```{r}
sleep[1, ]
```

## Question 7  
### We are interested in two groups; small and large mammals. Let's define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.  
```{r}
small_sleep <- 
  filter(sleep, bodywt <=1)

small_sleep_bodywtdataframe <- 
  select(small_sleep, bodywt)

small_sleep_bodywtdataframe
```

```{r}
big_sleep <- 
  filter(sleep, bodywt >=200)

big_sleep_bodywtdataframe <- 
  select(big_sleep, bodywt)

big_sleep_bodywtdataframe
```

## Question 8  
### What is the mean weight for both the large and small mammals?  
```{r}
summary(small_sleep$bodywt)
summary(big_sleep$bodywt)
```

## Question 9  
### Let's try to figure out if large mammals sleep, on average, longer than small mammals. What is the average sleep duration for large mammals as we have defined them?  
```{r}
summary(big_sleep$sleep_total)
```

## Question 10  
### What is the average sleep duration for small mammals as we have defined them?  
```{r}
summary(small_sleep$sleep_total) 
```

## Question 11  
### Which animals are the sleepiest? Which sleep least 18 hours per day?  
```{r}
sleepiest_sleep <- 
  filter(sleep, sleep_total >= 18)
sleepiest_sleep
```
The Little Brown Bat is the sleepiest animal!  

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  