---
title: "Lab 1 Homework"
subtitle: "01/17/20"
author: "Vincent Castillo"
date: "Winter 2020"
output:
  html_document:
    keep_md: TRUE
    theme: darkly
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run.  

1. Navigate to the R console and calculate the following expressions.  
  + 5 - 3 * 2  
  + 8 / 2 ** 2  
```{r}
+ 5 - 3 * 2  
  + 8 / 2 ** 2  
```

2. Did any of the results in #1 surprise you? Write two programs that calculate each expression such that the result for the first example is 4 and the second example is 16.  
```{r}
(5-3)*2
(8/2)**2
```

3. Make a new object `pi` as 3.14159265359. 
```{r}
pi <- 3.14159265359
```

## Question 4
```{r}
class(pi)
```

## Question 5
a)
```{r}
days <- c( "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
```

```{r}
blackjack<- c(140, -20, 70, -120, 240, NA, NA)
```

```{r}
roulette<- c(60, 50, 120, -300, 10, NA, NA)
```

b)
```{r}
sum(blackjack)
```
c)
```{r}
new_blackjack <- na.omit (blackjack)
```

```{r}
sum(new_blackjack)
```
d)
```{r}
new_roulette <- na.omit (roulette)
```

```{r}
sum(new_roulette)
```
e)
```{r}
total_week <- c(blackjack+roulette)
```

```{r}
total_week
```
f) Stick to blackjack
```{r}
blackjack>roulette
```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  